---
title: "Analysis of example trips for Surf Coast Shire"
author: Dhirendra Singh
date: 12 April, 2018
output:
  html_document:
    keep_md: yes
#  md_document:
#    variant: markdown_github
---

## Anglesea: `./trips-anglesea.csv.gz` 

```{r include=FALSE}
csv <- read.csv(gzfile("trips-anglesea.csv.gz"), header=T, sep=",")
```

Data columns:
```{r echo=FALSE}
paste(colnames(csv), collapse=",")
```

Number of unique persons:
```{r echo=FALSE} 
length(unique(csv$personID))
```

Number of unique activity types:
```{r echo=FALSE} 
length(unique(csv$actType))
```

Number of unique origin nodes:
```{r echo=FALSE} 
length(unique(csv$Node1))
```

Number of unique destination nodes:
```{r echo=FALSE} 
length(unique(csv$Node2))
```


Plot of all nodes (*red=origins, blue=destinations*):
````{r echo=FALSE, fig.width=10, fig.height=10}
# Convert the EPSG:32754 coords to latlon for plotting
library(proj4)
proj4string <- "+proj=utm +zone=54s +south +ellps=WGS84 +datum=WGS84 +units=m +no_defs "
xy <- unique(data.frame(x=csv$X, y=csv$Y))
pj <- project(xy, proj4string, inverse=TRUE)
latlon <- data.frame(lat=pj$y, lon=pj$x)
xy <- unique(data.frame(x=csv$DestX, y=csv$DestY))
pj <- project(xy, proj4string, inverse=TRUE)
latlonDest <- data.frame(lat=pj$y, lon=pj$x)
# plot the points on a map
suppressMessages({
  library(ggmap)
  map<-get_googlemap(
    center='-38.367370, 144.026623'
    , maptype="terrain"
    , size=c(640, 640)
    , scale=2
    , format="png8"
    , zoom=10
    )
  p<-ggmap(map)
  p + 
    geom_point(data=latlon, colour='red',size=1) + 
    geom_point(data=latlonDest, colour='blue',size=2)
  #show(p)
})
```

## Surf Coast Shire: `./trips-scs.csv.gz` 

```{r include=FALSE}
csv <- read.csv(gzfile("trips-scs.csv.gz"), header=T, sep=",")
```

Data columns:
```{r echo=FALSE}
paste(colnames(csv), collapse=",")
```

Number of unique persons:
```{r echo=FALSE} 
length(unique(csv$person.id))
```

Number of unique activity types:
```{r echo=FALSE} 
length(unique(csv$type))
```

Number of unique origin nodes:
```{r echo=FALSE} 
length(unique(csv$Node.no1))
```

Number of unique destination nodes:
```{r echo=FALSE} 
length(unique(csv$Node.no2))
```


Plot of all nodes (*red=origins, blue=destinations*):
````{r echo=FALSE, fig.width=10, fig.height=10}
# Convert the EPSG:32754 coords to latlon for plotting
#library(proj4)
proj4string <- "+proj=utm +zone=54s +south +ellps=WGS84 +datum=WGS84 +units=m +no_defs "
xy <- unique(data.frame(x=csv$X, y=csv$Y))
pj <- project(xy, proj4string, inverse=TRUE)
latlon <- data.frame(lat=pj$y, lon=pj$x)
xy <- unique(data.frame(x=csv$DestY, y=csv$DestX)) # header has values reversed X is Y and vice-versa!
pj <- project(xy, proj4string, inverse=TRUE)
latlonDest <- data.frame(lat=pj$y, lon=pj$x)

# plot the points on a map
suppressMessages({
  library(ggmap)
  map<-get_googlemap(
    center='-38.324116, 144.062375'
    , maptype="terrain"
    , size=c(640, 640)
    , scale=2
    , format="png8"
    , zoom=10
    )
  p<-ggmap(map)
  p + 
    geom_point(data=latlon, colour='red',size=2) + 
    geom_point(data=latlonDest, colour='blue',size=1)
  #show(p)
})
```